---
title: "Heart Disease Classification - Revisited"
author: "Team 3"
date: '2023-07-03'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

0. Import relevant libraries

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
library(ROCR)
library(irr)
library(C50)
library(caret)
library(e1071)
library(ranger)
library(parallel)
library(doParallel)
library(pROC)
library(readxl)
library(rpart)
library(rpart.plot)
library(e1071)
library(reshape2)

#Set global options to nonscientific notation
options(scipen = 999)

```

1. Convert XPT file to csv

```{r}

#This is only done once so I will comment it out 
#Library for converting .XPT to .CSV
#library(foreign)

#Read in .XPT file
#heart <- read.xport("/Users/paige/temp/heartDiseaseDM/LLCP2021.XPT ")

#Write to .CSV
#write.csv(file = "heartData2021.csv", heart, row.names=FALSE)

```

2. Import data set

```{r}

#Read in .CSV file
heart <- read.csv("heartData2021.csv")
heart2 <- read.csv("heartData2021.csv")
```

3. Data pre-processing

a)  Initial dimensionality reduction
- Removed all non-core questions (irrelevant attributes)
- Removed redundant attributes (calculations from other variables, other than BMI and age)
- Removed any introductory questions (irrelevant attributes)
- Removed any variables who had greater than 50% of the values missing 

```{r}

#Remove all non-core questions
#Exception X_AGE_G because there is not a non-calculated field that displays raw aged. These are imputed in six groups.
#Exception X_BMI5CAT to replace weight/height rather than manually calculating
#Exception X_MICHD because CDC has combined CVDINFR4 (myocardial infarction) and CVDCRHD4 (coronary heart disease) - target variable
heart <- heart[,c(1:97,247,265,270)]

#Remove columns that are introductory and calculated fields
heart <- heart[,-c(2:31)]

#Remove height and weight variables since BMI will be used instead
heart <- heart %>% select(-c(WEIGHT2, HEIGHT3))

#CVDINFR4 (myocardial infarction) and CVDCRHD4 (coronary heart disease) since these are a derivation of X_MICHD
heart <- heart %>% select(-c(CVDINFR4, CVDCRHD4))

#Calculate proportion of NA values by variable
naPercentages <- sort((colMeans(is.na(heart))) * 100, decreasing = TRUE)

#Select all that have a proportion missing > 50
lowProportionVars <- names(which(naPercentages > 50))

#Remove high proportion missing from dataset
heart <- heart[, !(names(heart) %in% lowProportionVars)]

#Remove rows that have target class missing
heart <- heart[complete.cases(heart$X_MICHD), ]

str(heart)
```
b) Data cleaning 

```{r}

#Convert everything but numeric variables to character
numericVars <- c("POORHLTH","PHYSHLTH","MENTHLTH","CPDEMO1B","ALCDAY5")

heart <- heart %>%
  mutate_at(vars(-one_of(numericVars)), as.character)

str(heart)

```

The majority of the features in this data set are coded responses and will need to be converted to their code descriptions prior so it is easier to interpret findings. Since there are so many variables to change I have used the file Format21.sas to match the descriptions to the codes in mass.

Cleanup of code file:

```{r}

#Read in SAS code dataset
codeFile <- read_excel("heartDiseaseSASCodes.xlsx", sheet = "Sheet1")

#Rename column
colnames(codeFile)[1] <- "Column"

#Create a new column to store the values starting with a letter (the variable name rows)
codeFile$varNames <- NA

#Iterate over each row in the data frame
for (i in 1:nrow(codeFile)) {
  
  #Convert value to character for regex
  value <- as.character(codeFile$Column[i])
  
  #Check if the value starts with a letter
  if (grepl("^[A-Za-z]", value)) {
    #If starts with a letter, then move the variable to a new column
    codeFile$varNames[i] <- value
    codeFile$Column[i] <- NA
  }
}

#Fill in variable names in spaces between two variables
codeFile <- fill(codeFile, varNames)

#Remove any rows with NA
codeFile <- codeFile[complete.cases(codeFile), ]

#Subset data to remove any rows with ; 
codeFile <- subset(codeFile, Column != ";")

#Split the column into two columns according to the = sign
splitColumn <- strsplit(codeFile$Column, "=")

# Creating separate columns for values and descriptions
codeFile$codeVal <- sapply(splitColumn, "[", 1)
codeFile$codeDesc <- sapply(splitColumn, "[", 2)

#Remove " from code description
codeFile$codeDesc <- gsub("\"", "", codeFile$codeDesc)

#Remove . from code value
codeFile$codeVal <- gsub("\\.", "", codeFile$codeVal)

#Remove spaces from code value
codeFile$codeVal <- gsub("\\s", "", codeFile$codeVal)

#Remove trailing white space from code description
codeFile$codeDesc <- trimws(codeFile$codeDesc)

#Select only data that has numbers as code value
codeFile <- subset(codeFile, !(codeVal %in% c("D", "R", "")))

#Remove original column
codeFile <- codeFile[,-1]

#Some of the variable names in the code file do not match what's in the dataset
#I renamed them manually
codeFile$varNames <- recode(codeFile$varNames,"AD3DEPEV"="ADDEPEV3", "ALC5DAY"="ALCDAY5", "ALC5DAY"="ALCDAY5", "AST3HMA"="ASTHMA3", "BP6HIGH"="BPHIGH6", "CHC3COPD"="CHCCOPD3", "CHC2KDNY"="CHCKDNY2", "CHECK1UP"="CHECKUP1", "CHOL3CHK"="CHOLCHK3", "CHOL3MED"="CHOLMED3", "CP1DEMO"="CPDEMO1B", "CVD4CRHD"="CVDCRHD4", "CVD4INFR"="CVDINFR4", "CVD3STRK"="CVDSTRK3", "DIA4BETE"="DIABETE4", "ECIG1NOW"="ECIGNOW1", "EMPLOY1F"="EMPLOY1", "EXER2ANY"="EXERANY2", "FLU7SHOT"="FLUSHOT7", "FRENCH1F"="FRENCHF1", "FRUIT2F"="FRUIT2", "FRUIT2JU"="FRUITJU2", "FV1GREEN"="FVGREEN1", "HAV5ARTH"="HAVARTH5", "HEIGHT"="HEIGHT3", "HIV7TST"="HIVTST7", "IN3COME"="INCOME3", "MED1COST"="MEDCOST1", "PERS3DOC"="PERSDOC3", "PNEU4VAC"="PNEUVAC4", "RENT1HOM"="RENTHOM1", "SMOK100_"="SMOKE100", "TOLD3HI"="TOLDHI3", "USE3NOW"="USENOW3", "VEGE2TAB"="VEGETAB2", "VET3ERAN"="VETERAN3", "WEIGHT"="WEIGHT2")

```

Recode all categorical coded responses in the heart dataset: 

```{r}
#Object for all variable names in reduced heart dataset
allValues <- as.list(names(heart))

#Remove the numerical variables from the coding 
allValues <- setdiff(allValues, numericVars)

#Loop through each variable name in the heart dataset
for (var in allValues) {
  #Loop through each row of the codeFile dataset
  for (i in 1:nrow(codeFile)) {
    #Check if the value in the first column of the code file matches the variable of the heart dataset 
    if (codeFile[i, 1] == var) {
      # Output values from the second and third columns of the codefile (the code value and description)
      outputValue1 <- as.character(codeFile[i, 2])
      outputValue2 <- as.character(codeFile[i, 3])
      #Replace all coded values 
      heart[[var]][heart[[var]] == outputValue1] <- outputValue2
    }
  }
}

```

Manual recode of three calculated variables since these variables are not in the code file: 

```{r}

#Code X_AGE_G manually because this is a calculated variable and is not in the code file
heart$X_AGE_G <- recode(heart$X_AGE_G,"1" = "Age 18 to 24", "2" = "Age 25 to 34", "3" = "Age 35 to 44","4" = "Age 45 to 54", "5" = "Age 55 to 64","6" = "Age 65 or older")

#Code X_BMI5CAT manually because this is a calculated variable and is not in the code file
heart$X_BMI5CAT <- recode(heart$X_BMI5CAT,"1" = "Underweight", "2" = "Normal weight", "3" = "Overweight","4" = "Obese")

#Code X_MICHD manually because this is a calculated variable and is not in the code file
heart$X_MICHD <- recode(heart$X_MICHD,"1" = "Reported having MI or CHD", "2" = "Did not report having MI or CHD")

```

3. Imputation of missing values

Every feature with categorical values has been imputed using mode that is either:

A) Grouped by another feature if there is variation between groups age, sex, or region
B) Mode of entire feature if there is no noticeable variation between groups age, sex, or region

If numerical feature, either median or mean used depending on data distribution, grouped by either age, sex, region

```{r}

#mode function - will be used for imputation
getmode <- function(v) {
   uniqv <- unique(na.omit(v))
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

```

Discretization: New column added for regions of the US and state removed

```{r}

#X_STATE -  State FIPS Code
round(prop.table(table(heart$X_STATE, useNA = "always")) * 100, 2)

#New column for regions of US
heart$X_REGION <- ifelse(heart$X_STATE %in% c("9","23","25","33","44","50","34","36","42"),"Northeast",
                        ifelse(heart$X_STATE %in% c("17","18","26","39","55","19","20","27","29","31","38","46"), "Midwest",
                               ifelse(heart$X_STATE %in% c("10","13","24","37","45","51","11","54","1","21","28","47","5","22","40","48"), "South",
                                      ifelse(heart$X_STATE %in% c("4","8","16","30","32","35","49","56","2","6","15","41","53"), "West",
                                        "Territories"))))

round(prop.table(table(heart$X_REGION, useNA = "always")) * 100, digits = 1)

#Remove state variable
heart <- heart %>% select(-X_STATE)

```

Imputation of categorical features using mode:

Rules:
If the missing values are less than 1%, then just use the mode of the entire variable
If the missing values are >1%, use mode grouped by region, age, or gender, whichever shows any variance in mode. If none show variance then use the mode of the entire variable.
If more then one grouping shows variance, verify which one shows the most variance and use 

```{r}

#Select all categorical features
characterColumns <- heart %>% select_if(is.character)

#Set percentage of missing values
naPercentages <- sort((colMeans(is.na(characterColumns))) * 100, decreasing = TRUE)

#Variable with less than 1% missing
onePercent <- names(which(naPercentages < 1))

#Loop through every variable that has <1% missing and replace NA with mode
for (var in onePercent) {
  #Loop through each row of the codeFile dataset
  heart <- heart %>% mutate(!!var := ifelse(is.na(!!sym(var)),getmode(!!sym(var)), !!sym(var)))
}

#Select all categorical features
characterColumns <- heart %>% select_if(is.character)

#Calculate percentages of missing values
naPercentages <- sort((colMeans(is.na(characterColumns))) * 100, decreasing = TRUE)

#Everything that has missing values selected
toImpute <- names(which(naPercentages > 0))

#Create empty data frame to store the results
whatsLeft <- data.frame()

#Loop through every variable to find modes grouped by age, sex, region
for (var in toImpute) {

#Calculate modes grouped by age, sex, region
modeCalcAge <- heart %>% group_by(X_AGE_G) %>% summarise(mode = getmode(!!sym(var)))
modeCalcGender <- heart %>% group_by(SEXVAR) %>% summarise(mode = getmode(!!sym(var)))
modeCalcRegion <- heart %>% group_by(X_REGION) %>% summarise(mode = getmode(!!sym(var)))

#If the modes are all the same, then mark as "no variance", else "variance"
modeVarAge <- ifelse(length(unique(modeCalcAge$mode)) == 1,"No variance","Variance")
modeVarGender <- ifelse(length(unique(modeCalcGender$mode)) == 1,"No variance","Variance")
modeVarRegion <- ifelse(length(unique(modeCalcRegion$mode)) == 1,"No variance","Variance")

variances <- data.frame(Var = var, Age = modeVarAge, Gender = modeVarGender, Region = modeVarRegion)

#If all the modes show no variance, then use mode of the entire variable
if (modeVarAge == "No variance" && modeVarGender == "No variance" && modeVarRegion == "No variance") 
  {
    heart <- heart %>% mutate(!!var := ifelse(is.na(!!sym(var)),getmode(!!sym(var)), !!sym(var)))
  } 

#If age shows variance, use age
else if(modeVarAge == "Variance" && modeVarGender == "No variance" && modeVarRegion == "No variance")
  {
    heart <- heart %>% group_by(X_AGE_G) %>% mutate(!!var := ifelse(is.na(!!sym(var)),getmode(!!sym(var)), !!sym(var)))
}

#Add whatever is left to a data frame
else 
  {
    whatsLeft <- rbind(whatsLeft, variances)
  }
}

#Print data frame
print(whatsLeft)

#Manual correction of final three categorical variables
#FLUSHOT7
heart %>% group_by(X_AGE_G) %>% summarise(mode = getmode(FLUSHOT7))
heart %>% group_by(SEXVAR) %>% summarise(mode = getmode(FLUSHOT7))
heart %>% group_by(X_REGION) %>% summarise(mode = getmode(FLUSHOT7))
heart <- heart %>% group_by(X_AGE_G) %>% mutate(FLUSHOT7 = ifelse(is.na(FLUSHOT7),getmode(FLUSHOT7), FLUSHOT7))

#INCOME3
heart %>% group_by(X_AGE_G) %>% summarise(mode = getmode(INCOME3))
heart %>% group_by(SEXVAR) %>% summarise(mode = getmode(INCOME3))
heart %>% group_by(X_REGION) %>% summarise(mode = getmode(INCOME3))
heart <- heart %>% group_by(X_AGE_G) %>% mutate(INCOME3 = ifelse(is.na(INCOME3),getmode(INCOME3), INCOME3))

#X_BMI5CAT
heart %>% group_by(X_AGE_G) %>% summarise(mode = getmode(X_BMI5CAT))
heart %>% group_by(SEXVAR) %>% summarise(mode = getmode(X_BMI5CAT))
heart %>% group_by(X_REGION) %>% summarise(mode = getmode(X_BMI5CAT))
heart <- heart %>% group_by(X_AGE_G) %>% mutate(X_BMI5CAT = ifelse(is.na(X_BMI5CAT),getmode(X_BMI5CAT), X_BMI5CAT))
```

Recode numerical variables:
There are not many numerical variables but each have to be addresses differently because they have different codings.

```{r}

#POORHLTH, PHYSHLTH and MENTHLTH all have the same values 1-30, Refused, Don't know, or Blank
daysVariables <- c("POORHLTH","PHYSHLTH","MENTHLTH")

#Loop through three variable to replace with values
for (var in daysVariables) {
  heart <- heart %>% mutate(!!var := ifelse(!!sym(var) == 88, 0, ifelse(!!sym(var) == 77 | !!sym(var) == 99,NA,!!sym(var))))
}

#Verify conversion
table(heart$POORHLTH, useNA = "always")
table(heart$PHYSHLTH, useNA = "always")
table(heart$MENTHLTH, useNA = "always")

#CPDEMO1B - replace 7, 8, 9 with NA or 0
heart <- heart %>% mutate(CPDEMO1B = ifelse(CPDEMO1B == 8, 0, ifelse(CPDEMO1B == 7 | CPDEMO1B == 9, NA, CPDEMO1B)))

#Verify conversion
table(heart$CPDEMO1B, useNA = "always")

#Create a new variable called DRINKSYEARLY that calculates number of drinks per year based on ALCDAY5
heart <- heart %>% mutate(DRINKSYEARLY = ifelse(ALCDAY5 <= 107 & ALCDAY5 >= 101, (ALCDAY5 %% 100)*52,
                                           ifelse(ALCDAY5 <= 230 & ALCDAY5 >= 201, (ALCDAY5 %% 200)*12,
                                           ifelse(ALCDAY5 == 777 | ALCDAY5 == 999, NA, 
                                                  ifelse(ALCDAY5 == 888, 0, ALCDAY5)))))

#Remove ALCDAY5 since I made a calculated variable
heart <- heart %>% select(-ALCDAY5)
str(heart)
```

Impute numerical variables:

```{r}

medianMean <- function(var) {
print(heart %>% group_by(X_AGE_G) %>% summarise(mean = mean(!!sym(var), na.rm = TRUE), median = median(!!sym(var), na.rm = TRUE)))
print(heart %>% group_by(SEXVAR) %>% summarise(mean = mean(!!sym(var), na.rm = TRUE), median = median(!!sym(var), na.rm = TRUE)))
print(heart %>% group_by(X_REGION) %>% summarise(mean = mean(!!sym(var), na.rm = TRUE), median = median(!!sym(var), na.rm = TRUE)))
}

medianMean("POORHLTH")
medianMean("PHYSHLTH")
medianMean("MENTHLTH")
medianMean("DRINKSYEARLY")
medianMean("CPDEMO1B")

heart <- heart %>% group_by(X_AGE_G) %>% mutate(POORHLTH = ifelse(is.na(POORHLTH), round(mean(POORHLTH, na.rm = TRUE)), POORHLTH))
heart <- heart %>% group_by(X_AGE_G) %>% mutate(PHYSHLTH = ifelse(is.na(PHYSHLTH), round(mean(PHYSHLTH, na.rm = TRUE)), PHYSHLTH))
heart <- heart %>% group_by(X_AGE_G) %>% mutate(MENTHLTH = ifelse(is.na(MENTHLTH), round(mean(MENTHLTH, na.rm = TRUE)), MENTHLTH))
heart <- heart %>% group_by(X_AGE_G) %>% mutate(DRINKSYEARLY = ifelse(is.na(DRINKSYEARLY), median(DRINKSYEARLY, na.rm = TRUE), DRINKSYEARLY))
heart <- heart %>% mutate(CPDEMO1B = ifelse(is.na(CPDEMO1B),round(mean(CPDEMO1B, na.rm = TRUE)), CPDEMO1B))

```

Verify no missing values:

```{r}
#Count number of rows in dataset
nrow(heart)

#Count number of rows without missing values
nrow(heart[complete.cases(heart),])

```

4. Dimensionality Reduction Using Decision Tree

```{r}
#Turn characters into factors
heartModeling <- heart %>% mutate_if(is.character, as.factor)
heartModeling$X_AGE_G <- as.factor(heartModeling$X_AGE_G)

#Verify changes
str(heartModeling)

#Check proportion of X_MICHD
prop.table(table(heartModeling$X_MICHD))

#Build the simplest decision tree splitting target class 50/50
heartDTModelSimple <- rpart(X_MICHD ~ ., data = heartModeling, parms=list(prior=c(.5,.5)), method = 'class')

#Display tree
rpart.plot(heartDTModelSimple)
rpart.plot(heartDTModelSimple, text = TRUE, fallen.leaves = TRUE, fallen.col = "white", fallen.text.col.fill = "black")

#View summary of model
heartDTModelSimpleSummary <- summary(heartDTModelSimple)

#Select variables of importance in decision tree
modelColumns <- names(heartDTModelSimpleSummary$variable.importance)

#Add on the target variables 
modelColumns <- c(modelColumns,"X_MICHD")

#Create new data frame of selected variables for training models
heartModelSmall <- heartModeling[, names(heartModeling) %in% modelColumns]

```

4. Split data into training and test sets

```{r}

#Number of test set 25%
rowN <- nrow(heartModelSmall)
testRowN <- rowN*.25

#Sample numbers for test set
set.seed(123)
heartSample <- sample(rowN, testRowN)

#25% Test Set 
heartTest <- heartModelSmall[heartSample,]

#75% Training Set
heartTrain <- heartModelSmall[-heartSample,]

#Verify proportions
prop.table(table(heartTest$X_MICHD))
prop.table(table(heartTrain$X_MICHD))

```

5. Modeling

Model 1: Naive Bayes 

```{r}

heartModel_1 <- naiveBayes(X_MICHD ~ .,
                           data = heartTrain)

#Evaluate model performance
heartPred_1 <- predict(heartModel_1, heartTest)

#Confusion matrix 
results_1 <- confusionMatrix(data=heartPred_1,
                             reference = heartTest$X_MICHD,
                             positive = "Reported having MI or CHD")

#Showing only Kappa
results_1$overall['Kappa']

```

Model 2: Random Forest 

```{r}

ctrl <- trainControl(method = "cv", sampling = "up", selectionFunction = "oneSE", savePredictions = T)
ctrl

gridRf <- expand.grid(mtry = 20, splitrule = 'hellinger', min.node.size = 1)
gridRf

set.seed(300)
heartModel_2L <- train(X_MICHD ~ .,
                        data = heartTrain,
                        method = "ranger",
                        metric = "Kappa",
                        tuneGrid = gridRf,
                        trControl = ctrl)

#Display model
heartModel_2L

#Evaluate model performance
heartPred_2L <- predict(heartModel_2L, heartTest)

#Confusion matrix 
results_2L <- confusionMatrix(data=heartPred_2L,
                             reference = heartTest$X_MICHD,
                             positive = "Reported having MI or CHD")

results_2L

#Showing only kappa
results_2L$overall['Kappa']

```

Model 3: Bagging

```{r}

ctrl <- trainControl(method = "cv", sampling = "up", selectionFunction = "oneSE", savePredictions = T)

#cntrl grid
gridBag <- expand.grid(vars = seq(2,16, by = 2))
gridBag

bagCtrl <- bagControl(fit = ldaBag$fit,
                      predict = ldaBag$pred,
                      aggregate = ldaBag$aggregate)
set.seed(300)
heartModel_3 <- train(X_MICHD ~ .,
                        data = heartTrain,
                        method = "bag",
                        metric = "Kappa",
                        tuneGrid = gridBag,
                        bagControl = bagCtrl,
                        trControl = ctrl)

#Display model
heartModel_3

#Evaluate model performance
heartPred_3 <- predict(heartModel_3, heartTest)

#Confusion matrix 
results_3 <- confusionMatrix(data=heartPred_3,
                             reference = heartTest$X_MICHD,
                             positive = "Reported having MI or CHD")

results_3

#Showing only kappa
results_3$overall['Kappa']

```

Model 4: Logistic Regression 

```{r}
ctrl <- trainControl(method = "cv", sampling = "up", selectionFunction = "oneSE", savePredictions = T)

#cntrl grid
gridLog <- expand.grid(nIter = c(50,100,150,200))
gridLog

set.seed(300)
heartModel_4 <- train(X_MICHD ~ .,
                        data = heartTrain,
                        method = "LogitBoost",
                        metric = "Kappa",
                        tuneGrid = gridLog,
                        trControl = ctrl)

#Display model
heartModel_4

#Evaluate model performance
heartPred_4 <- predict(heartModel_4, heartTest)

#Confusion matrix 
results_4 <- confusionMatrix(data=heartPred_4,
                             reference = heartTest$X_MICHD,
                             positive = "Reported having MI or CHD")

results_4

#Showing only kappa
results_4$overall['Kappa']

```


Model 5: Decision Tree

```{r}

ctrl <- trainControl(method = "cv", sampling = "up", selectionFunction = "oneSE", savePredictions = T)

#cntrl grid
gridTree <- expand.grid(.model = "tree", .trials = c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19, 21), .winnow = TRUE)
gridTree



set.seed(300)
heartModel_5 <- train(X_MICHD ~ .,
                        data = heartTrain,
                        method = "C5.0",
                        metric = "Kappa",
                        tuneGrid = gridTree,
                        trControl = ctrl)

#Display model
heartModel_5

#Evaluate model performance
heartPred_5 <- predict(heartModel_5, heartTest)

#Confusion matrix 
results_5 <- confusionMatrix(data=heartPred_5,
                             reference = heartTest$X_MICHD,
                             positive = "Reported having MI or CHD")

results_5

#Showing only kappa

results_5$overall['Kappa']

```

6. Plots

a. Heart disease by age plot

```{r}

heartDiseaseVsAge <- ggplot(data = heart, aes(x = X_AGE_G)) + 
  geom_bar(aes(y = (..count..)/sum(..count..), fill = X_MICHD), position = 'stack') + 
  labs(x = 'Age', y = "Percentage of Respondents", title = 'Heart Disease - Distribution by Age') + 
  scale_fill_manual(name = "Heart Disease", values = c('Did not report having MI or CHD' = 'pink','Reported having MI or CHD' = 'red', na.value = 'grey')) + 
  scale_y_continuous(labels = scales::percent, breaks = seq(0,.4, by=  .05)) +
  theme(aspect.ratio = 4/3, axis.text.x = element_text(angle = 25), legend.position = c(.4,.85),
        legend.title = element_text(size = 6), legend.text = element_text(size = 6)) 
heartDiseaseVsAge

```

b. Heart disease by region

```{r}
#Heart Disease by Region
heartDiseaseVsRegion <- ggplot(data = heart, aes(x = X_REGION)) + 
  geom_bar(aes(y = (..count..)/sum(..count..), fill = X_MICHD), position = 'stack') + 
  labs(x = 'Region', y = "Percentage of Respondents", title = 'Distribution by Region') + 
  scale_fill_manual(name = "Heart Disease", values = c('Did not report having MI or CHD' = 'pink','Reported having MI or CHD' = 'red', na.value = 'grey')) +  
  scale_y_continuous(labels = scales::percent, breaks = seq(0,.4, by=  .05)) +
  theme(aspect.ratio = 4/3, axis.text.x = element_text(angle = 30), legend.position = c(.4,.85),
        legend.title = element_text(size = 6), legend.text = element_text(size = 6)) 
heartDiseaseVsRegion

```

c. Model evaluation plots

```{r}

#Model names for plotting
modelNames <- c('Naive Bayes', 'Random Forest', 'Bagging', 'Boosted Logistic Regression', 'Decision Tree')

#Kappa values for each model
kappaVals <- c(results_1$overall['Kappa'],results_2L$overall['Kappa'],results_3$overall['Kappa'],results_4$overall['Kappa'],results_5$overall['Kappa'])

#Accuracy values for each model
accuracyVals <-c(results_1$overall['Accuracy'],results_2L$overall['Accuracy'],results_3$overall['Accuracy'],results_4$overall['Accuracy'],results_5$overall['Accuracy'])

#Sensitivity values for each model
sensitivityVals <-c(results_1$byClass['Sensitivity'],results_2L$byClass['Sensitivity'],results_3$byClass['Sensitivity'],results_4$byClass['Sensitivity'],results_5$byClass['Sensitivity'])

#Combined all three measurements
modelDF <- data.frame(cbind(modelNames, kappaVals, accuracyVals, sensitivityVals))

```

```{r}

par(mfrow = c(3,1))
kappavsmodel <- ggplot() +
  geom_point(data = modelDFOld, aes(x = modelNames, y = kappaValsOld, color = "Initial"), size = 6) +
  geom_point(data = modelDF, aes(x = modelNames, y = kappaVals, color = "Enhanced"), size = 6) +
  labs(x = 'Model Types', y = 'Kappa Value', title = "Kappa Values - Initial vs. Enhanced") +
  scale_color_manual(values = c("Enhanced" = "pink", "Initial" = "red")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + 
  theme(axis.text.x = element_text(size = 10), legend.position="bottom", legend.title = element_blank()) + 
  scale_y_discrete(labels = function(x) sprintf("%.2f", as.numeric(x)))
kappavsmodel

#Accuracy model
accuracyvsmodel <- ggplot() +
  geom_point(data = modelDFOld, aes(x = modelNames, y = accuracyValsOld, color = "Initial"), size = 6) +
  geom_point(data = modelDF, aes(x = modelNames, y = accuracyVals, color = "Enhanced"), size = 6) +
  labs(x = 'Model Types', y = 'Accuracy Value', title = "Accuracy Values - Initial vs. Enhanced") +
  scale_color_manual(values = c("Enhanced" = "pink", "Initial" = "red")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + 
  theme(axis.text.x = element_text(size = 10), legend.position="bottom", legend.title = element_blank()) +
  scale_y_discrete(labels = function(x) sprintf("%.2f", as.numeric(x)))
accuracyvsmodel

#Sensitivity model
sensitivityvsmodel <- ggplot() +
  geom_point(data = modelDFOld, aes(x = modelNames, y = sensitivityValsOld, color = "Initial"), size = 6) +
  geom_point(data = modelDF, aes(x = modelNames, y = sensitivityVals, color = "Enhanced"), size = 6) +
  labs(x = 'Model Types', y = 'Sensitivity Value', title = "Sensitivity Values - Initial vs. Enhanced") +
  scale_color_manual(values = c("Enhanced" = "pink", "Initial" = "red")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + 
  theme(axis.text.x = element_text(size = 10), legend.position="bottom", legend.title = element_blank()) + 
  scale_y_discrete(labels = function(x) sprintf("%.2f", as.numeric(x)))
sensitivityvsmodel

combinedPlots <- grid.arrange(kappavsmodel, accuracyvsmodel, sensitivityvsmodel, ncol=3)

```

```{r}

#Model Effectiveness Plot
modelDF.m <- melt(modelDF, id.vars = 'modelNames')
modelDF.m$variable = rep(c('Kappa','Accuracy','Sensitivity'),each = 5)
ggplot(modelDF.m, aes(x = modelNames,y=value )) + 
  geom_point(size = 5, aes(shape = variable, color = variable)) + 
  labs(x = 'Model Types', y = 'Accuracy,Kappa,Sensitivity', title = "Model Effectiveness") +
  theme(axis.text.x = element_text(angle = 60)) 

```


```{r oldValues}

kappaValsOld <- c(0.1993979,0.1859396,0.2099951,0.3337035,0.2157214)
accuracyValsOld <- c(0.7004823,0.8366964,0.7269975,0.8299875,0.8254812)
sensitivityValsOld <- c(0.8142306,0.3753972,0.7675897,0.8306998,0.4681117)

modelDFOld <- data.frame(cbind(modelNames,kappaValsOld,accuracyValsOld,sensitivityValsOld))

#Model Effectiveness Plot
modelDF.mOld <- melt(modelDFOld, id.vars = 'modelNames')
modelDF.mOld$variable = rep(c('Kappa','Accuracy','Sensitivity'),each = 5)
ggplot(modelDF.mOld, aes(x = modelNames,y=value )) + geom_point(size = 4,  aes(shape = variable, color = variable)) + labs(x = 'Model Types', y = 'Accuracy,Kappa,Sensitivity', title = "Model Effectiveness") 

modelDFOldPlot <- ggplot() +
  geom_point(data = modelDFOld, aes(x = modelNames, y = sensitivityValsOld, color = "Sensitivity"), size = 6) +
  geom_point(data = modelDFOld, aes(x = modelNames, y = kappaValsOld, color = "Kappa"), size = 6) +
  geom_point(data = modelDFOld, aes(x = modelNames, y = accuracyValsOld, color = "Accuracy"), size = 6) +
  labs(x = 'Model Types', y = 'Value', title = "Original Modeling Results") +
  scale_color_manual(values = c("Sensitivity" = "pink", "Kappa" = "red", "Accuracy" = "grey")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + 
  theme(axis.text.x = element_text(size = 10), legend.position="bottom", legend.title = element_blank()) + 
  scale_y_discrete(labels = function(x) sprintf("%.2f", as.numeric(x)))
modelDFOldPlot

```





